---
title: "STAT 184 Final Project Preliminary EDA" 
author: "Abhishek Raj, Sarthak Batra and Parthiv Khaund"
output: html_notebook
---


## Final Project 


## Front Matter
```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)
library(leaflet)
library(DataComputing)
```


#### Loading Data

```{r}
crime_stats <- read.csv("crime.csv")

```


#### Summary Statistics
```{r}
str(crime_stats)
str(ZipGeography)
```

```{r}
summary(crime_stats)
```


```{r}
crime_stats
```

```{r}
Test<-
  crime_stats%>%
  select(OFFENSE_CODE_GROUP)%>%
  count(OFFENSE_CODE_GROUP, sort = TRUE)
Test
```

```{r}
Test1<-
  crime_stats%>%
  select(OFFENSE_CODE_GROUP,OFFENSE_DESCRIPTION)%>%
  filter(OFFENSE_CODE_GROUP == "Auto Theft")
Test1
```

```{r}
# top 10 criminal districs 
distric_count<-
  crime_stats %>%
  group_by(DISTRICT)%>%
  summarise(counter = n())%>%
  arrange(desc(counter))

distric_count <- distric_count[1:10,]
distric_top_10 <- distric_count[,1]
distric_top_10
```

```{r}
# creating a table of just lat and long 
lats<- 
  crime_stats%>%
  select(DISTRICT,Lat,Long)

map <- inner_join(distric_top_10,lats)
map <- map[1:10,]
```

```{r}
stationMap <-
  leaflet(map) %>%   # like ggplot()
  addTiles() %>%          # add the map
  addMarkers()

stationMap
```



#### Describe preliminary observations and intuition about the research question

Based on a preliminary observation from the graph we find that some months have more crime in a reported area.


### TWO Data Sources

#### Primary Data

The primary data that we will use is from https://www.kaggle.com/ankkur13/boston-crime-data 

#### Other data sources

ZipGeography is an another data source that we are planning to use.

