---
title: "STAT 184 Final Project Preliminary EDA" 
author: "Abhishek Raj, Sarthak Batra and Parthiv Khaund"
output: html_notebook
---


## Final Project 


## Front Matter
```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)
```


#### Loading Data

```{r}
crime_stats <- read.csv("crime.csv")
```


#### Summary Statistics
```{r}
str(crime_stats)
```

```{r}
summary(crime_stats)
```


```{r}
crime_stats
```

```{r}
Test<-
  crime_stats%>%
  select(OFFENSE_CODE_GROUP)%>%
  count(OFFENSE_CODE_GROUP, sort = TRUE)
Test
```

```{r}
Test1<-
  crime_stats%>%
  select(OFFENSE_CODE_GROUP,OFFENSE_DESCRIPTION)%>%
  filter(OFFENSE_CODE_GROUP == "Auto Theft")
Test1
```






















#### Describe preliminary observations and intuition about the research question

Based on a preliminary observation from the graph we find that some months have more crime in a reported area.


### TWO Data Sources

#### Primary Data

The primary data that we will use is from https://www.kaggle.com/ankkur13/boston-crime-data 

#### Other data sources

ZipGeography is an another data source that we are planning to use.